---
title: "Bar Charts"
author: "Alex Kong"
date: "11/3/2018"
output: html_document
---

Let's start off with the simple example as indicated on the EDAV website

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache=TRUE, tidy=FALSE)
```

```{r}
require(tidyverse)
require(reticulate)
path_to_python <- "/Library/Frameworks/Python.framework/Versions/3.7/bin/python3"
use_python(path_to_python, required=TRUE)
```

We're going to use HairEyeColor like the EDAV website to attempt to duplicate the graphs:

```{r}
colors <- as.data.frame(HairEyeColor)

# just female hair color, using dplyr
colors_female_hair <- colors %>%
  filter(Sex == "Female") %>%
  group_by(Hair) %>%
  summarise(Total = sum(Freq))

# take a look at data
head(colors_female_hair)

# write.csv(colors, file='hair_colors.csv') # more in the README about why I've done it this way
```

```{python}
import pandas as pd
import matplotlib.pyplot as plt
import seaborn as sns

colors = pd.read_csv('../data/hair_colors.csv')
colors.drop(columns='Unnamed: 0', inplace=True)
print(colors)
```

Here's a little Python data wrangling 101 for you.

```{python}
female_colors = colors.loc[colors.Sex == 'Female'].groupby(colors.Hair).sum().reset_index()
print(female_colors)
```

The bar chart is fairly easy to draw in Python. I love the colors!

```{python}
plt.clf()

_ = female_colors.plot.bar(x='Hair', y='Freq')
_ = plt.title('Bar Graph Using matplotlib in Python')
plt.legend().set_visible(False)
plt.show()
```

We can also flip the orientation.

```{python}
plt.clf()

_ = female_colors.plot.barh(x='Hair', y='Freq')
_ = plt.title('Bar Graph Using matplotlib in Python')
plt.legend().set_visible(False)
plt.show()
```

Keep in mind best visualization techniques.

```{python}
plt.clf()

female_colors.sort_values(by='Freq', ascending=False, inplace=True) # dealing with nominal categories, so sort by descending value
_ = female_colors.plot.bar(x='Hair', y='Freq')
_ = plt.title('Bar Graph Using matplotlib in Python')
plt.legend().set_visible(False)
plt.show()
```

Faceting in matplotlib is much harder than in ggplot. Here's one way to do it.

```{python}
plt.clf()

hair_colors = colors.Hair.unique()
fig, ax = plt.subplots(nrows=2, ncols=2)

for i, group in enumerate(hair_colors):
  colors.loc[colors.Hair == group].groupby(colors.Sex).sum().reset_index().plot.bar(x='Sex', y='Freq', ax=ax[i // 2][i % 2])

plt.show()
```

However, seaborn's FacetGrid IMO makes it much little easier.

```{python}
plt.clf()

fg = sns.FacetGrid(colors, col='Hair', col_wrap=2)
fg.map(sns.barplot, 'Sex', 'Freq', estimator=sum, ci=None)
plt.subplots_adjust(top=0.9)
fg.fig.suptitle('Faceted Seaborn barplot')
plt.show()
```

```{r}
require(datasets)
# write.csv(as.data.frame(Titanic), 'titanic.csv')
```

Let's try and generate the tl;dr barchart graph. It's not an easy task...

```{python}
plt.clf()

titanic = pd.read_csv('../data/titanic.csv').drop(columns='Unnamed: 0')
ship_grouped = titanic.groupby([titanic.Class, titanic.Sex, titanic.Survived]).sum().reset_index()
print(ship_grouped)

fg = sns.FacetGrid(ship_grouped, col='Class', col_wrap=2)
fg.map(sns.barplot, 'Survived', 'Freq', 'Sex', palette=['red', 'green'], estimator=sum, ci=None)

for ax in fg.axes:
  for p in ax.patches:
    ax.annotate("%.2f" % p.get_height(), (p.get_x() + p.get_width() / 2., p.get_height()),
                 ha='center', va='center', fontsize=11, color='gray', xytext=(0, 8),
                 textcoords='offset points') # taken from https://stackoverflow.com/questions/45946970/displaying-of-values-on-barchart

plt.legend().set_visible(True)
plt.show()
```

The preceding is absolutely not perfect and absolutely much easier to do in R!!!