---
title: "Histograms"
author: "Alex Kong (alk2225)"
date: "11/5/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache=TRUE, tidy=FALSE)
```

```{r}
require(tidyverse)
require(reticulate)
path_to_python <- "/Library/Frameworks/Python.framework/Versions/3.7/bin/python3"
use_python(path_to_python, required=TRUE)
```

Again, duplicate some of the graphs we saw on EDAV:

```{python}
import pandas as pd
import numpy as np
import matplotlib.pyplot as plt
import seaborn as sns
from scipy import stats

plt.clf()

x = [50, 51, 53, 55, 56, 60, 65, 65, 68]
plt.hist(x, bins=np.arange(50, 75, 5))
plt.xticks(np.arange(50, 75, 5))
plt.yticks(np.arange(0, 5, 1))
plt.show()
```

As you can see, histograms in matplotlib are default left closed. Right closed histograms are a little harder to get. Unfortunately, matplotlib's histogram function is built off of numpy's histogram function, which is default right closed. I haven't been able to find any easy way to change this yet, though using a simple hack, I was able to (kind of) duplicate right closed (I absolutely don't recommend this in real-life):

```{python}
plt.clf()

force_right_x = [meas - .1 if meas % 5 == 0 else meas for meas in x] # subtract a small value from all elements that fall on the boundary
print(force_right_x)

plt.hist(force_right_x, bins=np.arange(45, 75, 5))
plt.xticks(np.arange(50, 75, 5))
plt.yticks(np.arange(0, 5, 1))
plt.show()
```

As you can see, we can adjust the bin boundaries by passing a list of breaks into the bins parameter. This is much preferred over passing an integer value, which specifies the number of bins.

We can change the alignment by changing our boundaries:

```{python}
plt.clf()

plt.hist(x, bins=np.arange(50, 75, 5) - 2.5) # this centers the bins instead
plt.xticks(np.arange(50, 75, 5))
plt.yticks(np.arange(0, 5, 1))
plt.show()
```

Just for fun, we can also specify varied bin widths

```{python}
plt.clf()

plt.hist(x, bins=[45, 50, 60, 65, 70, 75])
plt.xticks(np.arange(50, 75, 5))
plt.yticks(np.arange(0, 5, 1))
plt.show()
```

By default, histograms are done using frequency counts. We can draw relative frequency histograms:

```{python}
plt.clf()

prices = [379, 425, 450, 450, 499, 529, 535, 535, 545, 599, 665, 675, 699, 699, 725, 725, 745, 799]
plt.hist(prices, bins=np.arange(300, 800, 100), weights=np.zeros_like(prices) + 1 / len(prices))
plt.show()
```

And density histograms:

```{python}
plt.clf()

plt.hist(prices, bins=np.arange(300, 800, 100), weights=np.zeros_like(prices) + 1 / (len(prices) * 100)) # 100 is the bin width
plt.yticks(np.arange(0, 0.004, 0.001))
plt.show()
```

Be careful interpreting a similar histogram with density=True, as it does not yield the same graph as shown in the lecture notes. It uses a different computational method, although the density overlay is still the same.

```{python}
plt.clf()

plt.hist(prices, bins=np.arange(300, 800, 100), density=True)
plt.yticks(np.arange(0, 0.004, 0.001))
plt.show()
```

We can also draw cumulative frequency histograms (though they're a little harder):

```{python}
plt.clf()

res = stats.cumfreq(prices, numbins=5, defaultreallimits=(300, 800))
plt.bar(np.arange(300, 800, 100) + 50, res.cumcount, width=res.binsize)
plt.yticks(np.arange(0, 18, 2))
plt.show()
```

Let's visualize the tl;dr graph on EDAV now:

```{r}
require(Sleuth3)
# write.csv(Sleuth3::case0201, 'finches.csv')
```

```{python}
plt.clf()

finches = pd.read_csv('../data/finches.csv')

fg = sns.FacetGrid(finches, col='Year', col_wrap=2)
fg.map(sns.distplot, 'Depth', norm_hist=True)

plt.subplots_adjust(top=0.7)
fg.fig.suptitle('Faceted histogram with density overlay')
plt.show()
```

As mentioned earlier, be careful interpreting the density bar values, as they are computed differently.