---
title: "Ridgeline Plots"
author: "Alex Kong (alk2225)"
date: "11/5/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache=TRUE, tidy=FALSE)
```

```{r}
require(tidyverse)
require(reticulate)
path_to_python <- "/Library/Frameworks/Python.framework/Versions/3.7/bin/python3"
use_python(path_to_python, required=TRUE)
```

```{r}
library(datasets)
# write.csv(Orange, '../data/orange.csv')
```

We're going to see if we can duplicate the results for the Orange dataset:

```{python}
from __future__ import unicode_literals # need to import this for joyplots to work
import pandas as pd
import numpy as np
import matplotlib.pyplot as plt
import seaborn as sns
from joypy import joyplot

orange = pd.read_csv('../data/orange.csv').drop(columns='Unnamed: 0')
print(orange)
```

Ridgeline plots in Python are implemented using joyplot from the joypy library:

```{python}
joyplot(orange, by='Tree', column='age', figsize=(5, 8))
plt.title('Joyplot example', pad=0.5)
plt.show()
```

We can further customize our joyplot:

```{python}
joyplot(orange, by='Tree', column='age', figsize=(5, 8), overlap=3, colormap=plt.cm.summer_r) # change the overlap and color scheme
plt.title('Joyplot example', pad=0.5)
plt.show()
```

Let's try the same on OrchardSprays:

```{r}
# write.csv(OrchardSprays, 'orchard_sprays.csv')
```

```{python}
orchard_sprays = pd.read_csv('../data/orchard_sprays.csv').drop(columns='Unnamed: 0')
joyplot(orchard_sprays, by='treatment', column='decrease', figsize=(5, 8), overlap=6, colormap=plt.cm.summer_r) # change the overlap and color scheme
plt.show()
```

As you can see, unlike geom_density_ridges, plots are ordered by increasing peak value from top to bottom.

Sadly, Python joyplots are fairly limited in their scope. An alternative would be to draw a bunch of density plots on top of each other using matplotlib or seaborn: https://stackoverflow.com/questions/17614499/frequency-trail-in-matplotlib.

However, joyplots can do this (the Unknown Pleasures covor inspiration):

```{python}
# technically, there's supposed to be no fill, but I really like the blue overlay for each plot
joyplot(orchard_sprays, by='treatment', column='decrease', figsize=(5, 8), kind='counts', background='k', linecolor='w', overlap=4)
plt.show()
```