---
title: "Scatterplots"
author: "Alex Kong (alk2225)"
date: "11/5/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache=TRUE, tidy=FALSE)
```

```{r}
require(tidyverse)
require(reticulate)
path_to_python <- "/Library/Frameworks/Python.framework/Versions/3.7/bin/python3"
use_python(path_to_python, required=TRUE)
```

```{python}
import pandas as pd
import numpy as np
import matplotlib.pyplot as plt
import seaborn as sns
```

Let's duplicate some of the EDAV scatterplot results:

```{python}
plt.clf()

speedski = pd.read_csv('speedski.csv').drop(columns='Unnamed: 0')
plt.scatter(speedski['Year'], speedski['Speed'])
plt.xlabel('Year', labelpad=0.4)
plt.ylabel('Speed', labelpad=0.01)
plt.show()
```

As with R scatterplots, consider alpha blending or jittering to prevent overlap.

```{python}
plt.clf()

wide_x = np.append(np.random.normal(loc=100, scale=2, size=50), np.random.normal(loc=10, scale=2, size=50)) # loc is mean, scale is standard deviation
wide_y = np.random.normal(loc=5, scale=2, size=100)

print(wide_x.tolist())
print(wide_y.tolist())

plt.scatter(wide_x.tolist(), wide_y.tolist())
plt.xlabel('x')
plt.ylabel('y')
plt.title('Example of scaling data in Python', pad=0.5)
plt.show()
```

Like R, we can add contour lines to our scatterplots (using Seaborn):

```{python}
sns.jointplot('Year', 'Speed', data=speedski, color='k').plot_joint(sns.kdeplot)
plt.show()
```

And we can plot scatterplot matrices as well:

```{python}
iris = pd.read_table('../data/iris.data', sep=',', names=['sepal_length', 'sepal_width', 'petal_length', 'petal_width', 'species'])
sns.pairplot(iris, hue="species")
plt.show()
```

The scatterplot matrix in seaborn is one of the only (if not the only) visualization techniques I prefer to use Python over R for due to its ease of use and prettier display IMO. Really a huge fan of the density curves, though histograms would also be acceptable.